Evolution of pre-copulatory and post-copulatory inbreeding behaviour

### 2016-09-26 ms_pre_post.Rmd
Revise first results section

We present our results by showing mean allele value changes over generations. Figures 1-5 show mean allele values of tendency for polyandry ($P_{a}$; red lines and shading), inbreeding mating strategy ($M_{a}$; blue lines and shading), and inbreeding fertilisation strategy ($F_{a}$; black lines and shading) over 10000 generations as calculated across all individuals (female and male) within a focal population over 40 replicate simulations with identical starting conditions. Solid lines represent the grand mean of replicate mean allele values, and shading around lines shows standard errors of mean allele values around the grand mean. Panels within figures show different cost combinations for tendency for polyandry ($c_{P}$), inbreeding mating strategy ($c_{M}$), and inbreeding fertilisation strategy ($c_{F}$). Figures have fewer than eight panels if some costs are not relevant because allele values are fixed, and fixed allele values are shown as horizontal lines where line height is set to the fixed allele value. Supporting Information S1 presents correlations of mean allele values across replicate simulations, which were generally weak; we found no evidence suggesting strong evolutionary divergence among replicates with identical starting conditions (e.g., as might be the case if populations stochastically evolved toward inbreeding avoidance through mate choice or fertilisation, but not both), meaning that inspecting grand means of replicate allele values accurately reflects expected evolutionary dynamics over generations.

@@

We first test our hypothesis that a focal female's ability to bias fertilisation with respect to kinship causes selection for polyandry by comparing simulations in which tendency for polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy can all evolve to simulations in which inbreeding fertilisation strategy is fixed to random mating with respect to kinship, as presented in Figures 1 and 2, respectively. Comparison of Figures 1 and 2 shows that evolution of inbreeding fertilisation strategy facilitates evolution of polyandry, particularly when costs of tendency for polyandry and inbreeding fertilisation strategy are zero. For example, in Figure 2A, tendency for polyandry allele values consistently become negative over generations ($P_{a}<0$; decreasing red line), causing populations to be monandrous. Females in these populations avoid mating with kin ($M_{a}<0$; decreasing blue line), and because these females sample mates with replacement, males that remain after their initial mate choice are more likely to be relatives, causing selection for monandry because females are most likely to succeed in avoiding inbreeding in their initial mate choice [see also @Duthie]. However, Figure 1A shows that when a focal female can also bias the probability that a particular mate sires her offspring, inbreeding fertilisation strategy evolves to inbreeding avoidance ($F_{a}<0$; decreasing black line). Consequently, the ability to bias fertilisation to avoid inbreeding causes polyandry to become adaptive, and positive values of tendency for polyandry alleles increase in frequency ($P_{a}>0$; increasing red line of Figure 1A). Although non-zero costs decrease net selection for both tendency for polyandry and inbreeding fertilisation strategy (see below), across all combinations of costs, tendency for polyandry allele values are more positive when inbreeding fertilisation strategy can also evolve. Evolution of post-copulatory inbreeding avoidance therefore greatly facilitates evolution of polyandry.

@@

Selection given constraints on evolution
----------------------------------------------

Figure 3 shows evolution of tendency for polyandry and inbreeding fertilisation strategy when inbreeding mating strategy is fixed so that females must mate randomly (blue line at $M_{a}=0$). For all cost combinations of tendency for polyandry and inbreeding fertilisation strategy (Figure 3 panels A-D), we observed selection for positive values of tendency for polyandry alleles and negative values of inbreeding fertilisation strategy alleles, which we interpret as strong evidence for evolution of polyandry and associated inbreeding avoidance through fertiisation bias. This pattern of evolution contrasts with that observed when inbreeding mating strategy was not fixed, indicating that constraints on evolution of inbreeding mating strategy increased the strength of selection for evolution of polyandry and inbreeding avoidance through fertilisation bias. For example, in Figures 1C,D,G,H, tendency for polyandry allele values were typically less than zero, reflecting evolution of monandry, and inbreeding fertilisation strategy allele values were not as negative as they were in simulations with fixed random mating (Figure 3). In general, when females were unable to avoid mating with relatives given severe inbreeding depression ($\beta=3$), selection for evolution of polyandry and post-copulatory inbreeding avoidance was strengthened.

Figure 2, which we previously contrasted with Figure 1 to test whether or not evolution of inbreeding fertilisation strategy facilitates selection for polyandry, shows evolution of tendency for polyandry and inbreeding mating strategy when inbreeding fertilisation strategy is fixed such that the sire of a focal female's offspring is randomly assigned from her mates according to a fair raffle. In all cost combinations of Figure 2 (panels A-D), tendency for polyandry allele values decreased over generations, reflecting evolution of monandry. However, values of inbreeding mating strategy alleles did not greatly differ over generations when inbreeding fertilisation strategy was fixed, meaning that the effect of fixed inbreeding fertilisation strategy on increasing selection for inbreeding avoidance through mate choice was weak or nelgigible. A weak effect is discernable in the contrast between Figure 2B and Figure 1B, in which the grand mean inbreeding mating strategy allele value is $-0.627$ when inbreeding fertilisation strategy is not fixed (Figure 1B) versus $-1.036$ when it is fixed (Figure 2B) in generation 10000. Because females have 20 alleles with additive effects on inbreeding mating strategy, an average female was therefore expected to have mean inbreeding mating strategy phenotype values of $M_{p} = -0.627 \times 20 = -12.54$ and $M_{p} = -1.036 \times 20 = -20.72$ for not fixed versus fixed inbreeding fertilisation strategy, respectively. These values would cause an outbred focal female to be $2.135$ and $4.180$ times more likely to choose a non-relative than a full sibling as a mate given not fixed versus fixed inbreeding fertilisation strategy, respectively. Therefore, when females were unable to bias fertilisation to avoid inbreeding given severe inbreeding depression ($\beta=3$), selection for evolution of pre-copulatory inbreeding avoidance was strengthened.

>>>>

<!--- We present our results by showing mean allele value changes over generations. Figures 1-5 show mean allele values of tendency for polyandry ($P_{a}$; red lines and shading), inbreeding mating strategy ($M_{a}$; blue lines and shading), and inbreeding fertilisation strategy ($F_{a}$; black lines and shading) over 10000 generations as calculated across all individuals (female and male) within a focal population over 40 replicate simulations with identical starting conditions. Solid lines represent the grand mean of replicate mean allele values, and shading around lines shows standard errors of mean allele values around the grand mean. Panels within figures show different cost combinations for tendency for polyandry ($c_{P}$), inbreeding mating strategy ($c_{M}$), and inbreeding fertilisation strategy ($c_{F}$). Figures have fewer than eight panels if some costs are not relevant because allele values are fixed, and fixed allele values are shown as horizontal lines where line height is set to the fixed allele value. Supporting Information S1 presents correlations of mean allele values across replicate simulations, which were generally weak; we found no evidence suggesting strong evolutionary divergence among replicates with identical starting conditions (e.g., as might be the case if populations stochastically evolved toward inbreeding avoidance through mate choice or fertilisation, but not both), meaning that inspecting grand means of replicate allele values accurately reflects expected evolutionary dynamics over generations. --->

@@

We first test our hypothesis that a focal female's ability to express post-copulatory choice with respect to kinship causes selection for polyandry by comparing simulations in which post-copulatory inbreeding strategy is fixed to random mating with respect to kinship to simulations in which tendency for polyandry, pre-copulatory inbreeding strategy, and post-copulatory inbreeding strategy can all evolve. The left and right columns of Figure 1 show results for fixed versus evolving post-copulatory choice, respectively, across four rows of costs of tendency for polyandry. For low costs of tendency for polyandry, $c_{P} < 0.005$, selection caused expected tendency for polyandry allele values ($P_{a}$) to be positive only when post-copulatory inbreeding avoidance evolved (Figure 1A-D), and for all $c_{P}$, $P_{a}$ values were higher when post-copulatory inbreeding avoidance evolved than when it could not evolve. When post-copulatory inbreeding avoidance could not evolve, $P_{a}$ values were negative even in the absence of a direct cost ($c_{P}=0$; Figure 1A) because females that expressed polyandry were more likely to increase their inbreeding due to sampling males without replacement [@Duthie]. Strong post-copulatory inbreeding avoidance consistently evolved in all simulations, even when $P_{a}$ values were not expected to be positive (Figure 1F,H), because of the threshold nature of the polyandry phenotype; Figure 2 shows how phenotypic polyandry was common even when mean polyandry allele values within a populatoin were expected to be negative. In all simulations, pre-copulatory inbreeding avoidance also evolved ($M_{a}<0$), meaning that both pre- and post-copulatory inbreeding avoidance was predicted, and evolution of adaptive polyandry occurred given sufficiently low costs of tendency for polyandry.

<!--- Comparison of Figures 1 and 2 shows that evolution of inbreeding fertilisation strategy facilitates evolution of polyandry, particularly when costs of tendency for polyandry and inbreeding fertilisation strategy are zero. For example, in Figure 2A, tendency for polyandry allele values consistently become negative over generations ($P_{a}<0$; decreasing red line), causing populations to be monandrous. Females in these populations avoid mating with kin ($M_{a}<0$; decreasing blue line), and because these females sample mates with replacement, males that remain after their initial mate choice are more likely to be relatives, causing selection for monandry because females are most likely to succeed in avoiding inbreeding in their initial mate choice [see also @Duthie]. However, Figure 1A shows that when a focal female can also bias the probability that a particular mate sires her offspring, inbreeding fertilisation strategy evolves to inbreeding avoidance ($F_{a}<0$; decreasing black line). Consequently, the ability to bias fertilisation to avoid inbreeding causes polyandry to become adaptive, and positive values of tendency for polyandry alleles increase in frequency ($P_{a}>0$; increasing red line of Figure 1A). Although non-zero costs decrease net selection for both tendency for polyandry and inbreeding fertilisation strategy (see below), across all combinations of costs, tendency for polyandry allele values are more positive when inbreeding fertilisation strategy can also evolve. Evolution of post-copulatory inbreeding avoidance therefore greatly facilitates evolution of polyandry. --->

@@

<!--- Selection given constraints on evolution --->
<!--- ---------------------------------------------- --->

<!--- Figure 3 shows evolution of tendency for polyandry and inbreeding fertilisation strategy when inbreeding mating strategy is fixed so that females must mate randomly (blue line at $M_{a}=0$). For all cost combinations of tendency for polyandry and inbreeding fertilisation strategy (Figure 3 panels A-D), we observed selection for positive values of tendency for polyandry alleles and negative values of inbreeding fertilisation strategy alleles, which we interpret as strong evidence for evolution of polyandry and associated inbreeding avoidance through fertiisation bias. This pattern of evolution contrasts with that observed when inbreeding mating strategy was not fixed, indicating that constraints on evolution of inbreeding mating strategy increased the strength of selection for evolution of polyandry and inbreeding avoidance through fertilisation bias. For example, in Figures 1C,D,G,H, tendency for polyandry allele values were typically less than zero, reflecting evolution of monandry, and inbreeding fertilisation strategy allele values were not as negative as they were in simulations with fixed random mating (Figure 3). In general, when females were unable to avoid mating with relatives given severe inbreeding depression ($\beta=3$), selection for evolution of polyandry and post-copulatory inbreeding avoidance was strengthened. --->

<!--- Figure 2, which we previously contrasted with Figure 1 to test whether or not evolution of inbreeding fertilisation strategy facilitates selection for polyandry, shows evolution of tendency for polyandry and inbreeding mating strategy when inbreeding fertilisation strategy is fixed such that the sire of a focal female's offspring is randomly assigned from her mates according to a fair raffle. In all cost combinations of Figure 2 (panels A-D), tendency for polyandry allele values decreased over generations, reflecting evolution of monandry. However, values of inbreeding mating strategy alleles did not greatly differ over generations when inbreeding fertilisation strategy was fixed, meaning that the effect of fixed inbreeding fertilisation strategy on increasing selection for inbreeding avoidance through mate choice was weak or nelgigible. A weak effect is discernable in the contrast between Figure 2B and Figure 1B, in which the grand mean inbreeding mating strategy allele value is $-0.627$ when inbreeding fertilisation strategy is not fixed (Figure 1B) versus $-1.036$ when it is fixed (Figure 2B) in generation 10000. Because females have 20 alleles with additive effects on inbreeding mating strategy, an average female was therefore expected to have mean inbreeding mating strategy phenotype values of $M_{p} = -0.627 \times 20 = -12.54$ and $M_{p} = -1.036 \times 20 = -20.72$ for not fixed versus fixed inbreeding fertilisation strategy, respectively. These values would cause an outbred focal female to be $2.135$ and $4.180$ times more likely to choose a non-relative than a full sibling as a mate given not fixed versus fixed inbreeding fertilisation strategy, respectively. Therefore, when females were unable to bias fertilisation to avoid inbreeding given severe inbreeding depression ($\beta=3$), selection for evolution of pre-copulatory inbreeding avoidance was strengthened. --->
